### 2024-10-14 posts/BarthPapageorgeThom/BPT.qmd

---
title: "Genetic endowments and wealth inequality"
subtitle: |
  Daniel Barth, Nicholas W. Papageorge, Kevin Thom\
  Journal of Political Economy, 2020, vol. 128, No.4
author: "伊藤成朗"
engine: knitr
execute:
  echo: false
  freeze: auto
format:
  revealjs:
    standalone: true
    embed-resources: true
    slide-number: true
    incremental: true
    width: 1600
    height: 900
    preview-links: auto
    logo: GrootbergGiraffeHead.jpg
    footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/BPT.html>
    theme: [sky, sky-add.scss]
    linkcolor: red
    code-line-numbers: true
    code-copy: true
    code-block-border-left: "#31BAE9"
    toc: true
    toc-depth: 1
    template-partials: 
      - ../toc-slide.html
    include-in-header: ../sky_add_FilePointer.html
    include-after-body: ../toc-add.html
    toc-bg-image: "../../GrootbergGiraffe3.jpg"
lightbox: auto
filters:
  - parse-latex
  - imagify
imagify:
  header-includes: |
    \usetikzlibrary{calc}
    \usetikzlibrary{positioning}
    \usetikzlibrary{math}
    \usetikzlibrary{decorations.pathmorphing}
    \usetikzlibrary{shapes}
    \usetikzlibrary{arrows}
    \usetikzlibrary{backgrounds}
    \usetikzlibrary{fit}
    \usepackage{pgfplots}
  pdf-engine: lualatex
include-in-header:
  - text: |
      <style>
       .title {
        font-size: 2.0em;
      }
      </style>
#### From https://github.com/quarto-dev/quarto-cli/discussions/2951
#### I am happy with the left-indenting of the section line
#### Horizontal centering can be challenging (!) in quarto: stupid
#### https://github.com/quarto-dev/quarto-cli/issues/1231
#### It turns out an easy solution is :::{style="text-align: center;"}
  - text: |
      <style>
      .center-xy {
        margin: 0;
        position: absolute;
        top: 50%;
        left: 20%;
        -ms-transform: translateY(-50%), translateX(-50%);
        transform: translateY(-50%), translateX(-50%);
      }
      </style>
bibliography: c:/seiro/docs/seiro.bib
#### ../../../seiro.bib
resources:
  - BPT.pdf
# quarto render BPT.qmd
---
```{css, echo = F, results = "hide"}
Red {
  color: red;
}
Orange {
  color: #e94e2c;
}
Blue {
  color: blue;
}
.reveal .slides pre code {
    max-height: 800px;
}
#TOC {
  font-color: #e94e2c;
}
.big-code{
  font-size: 110%  
}
#### Below is suggested by Claude
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list ol {
####   list-style-type: none;
####   padding-left: 0;
####   margin-left: 0;
####   counter-reset: item;
#### }
#### .small-font-list ol li {
####   margin-bottom: 0.5em;
####   display: flex;
#### }
#### .small-font-list ol li::before {
####   content: counter(item) ". ";
####   counter-increment: item;
####   margin-right: 0.5em;
####   flex-shrink: 0;
#### }
#### .small-font-list ol li > p:first-child {
####   margin: 0;
####   display: inline;
#### }
.small-font-list {
  font-size: 70%;
  line-height: 1.2;
}
.small-font-list ol {
  list-style-type: none;
  padding-left: 0;
  margin-left: 0;
  counter-reset: item;
}
.small-font-list ol li {
  counter-increment: item;
  display: flex;
  align-items: flex-start;
  margin-bottom: 0.5em;
}
.small-font-list ol li::before {
  content: counter(item) ". ";
  flex: 0 0 2.5em;  /* Adjust this value to change the width of the number area */
  text-align: right;
  margin-right: 0.5em;
}
#### description list by Claude
/* Main container for the description list */
.description-list {
  font-size: 50% !important;  /* Reduces font size to 70% of the parent element */
  line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Styles for the description list element */
.description-list dl {
  margin-left: 0;  /* Removes default left margin */
  padding-left: 0;  /* Removes default left padding */
}
/* Styles for the term (header) in the description list */
.description-list dt {
  font-weight: bold;  /* Makes the term text bold */
  display: inline;  /* Displays the term inline. Uncomment for a line break after item header. */
  margin-right: 0.5em;  /* Adds some space after the term */
}
/* Styles for the description in the description list */
.description-list dd {
  display: inline;  /* Displays the description inline. Uncomment for a line break after item header. */
  margin-left: 0;  /* Removes default left margin */
}
/* Styles for paragraphs within the description */
.description-list dd p {
  display: inline;  /* Keeps paragraph content inline. Uncomment for a line break after item header. */
  margin: 0;  /* Removes default paragraph margins */
}
/* Creates a new line and adds space between list items */
.description-list dd::after {
  content: "";  /* Necessary for the pseudo-element to render */
  display: block;  /* Creates a line break */
  margin-bottom: 0.5em;  /* Adds space between items. Adjust as needed */
}
#### description with a line break between item header and texts
.description-lb {
  font-size: 70%;
  line-height: 1.2;
}
.description-lb dl {
  margin-left: 0;
  padding-left: 0;
}
.description-lb dt {
  font-weight: bold;
  margin-bottom: 0.2em;
}
.description-lb dd {
  margin-left: 0;
  margin-bottom: 1em;  /* Adjust this value to control spacing between items */
}
.description-lb dd p {
  margin: 0;
}
```

----

Genome

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
![Chromosone...DNA](DNAMacrostructure.jpg)

::: {style="font-size: 70%;"}
<https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication/>
:::
::::

::::{.fragment}
![DNA structure](DNAStructure.jpg)

::: {style="font-size: 70%;"}
<https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna->
:::
::::

:::::

----

<https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html>

<https://studylib.net/doc/5786293/snp-presentation>

----