### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

---
title: |
  | Genetic endowments
  | and wealth inequality
subtitle: |
  Daniel Barth, Nicholas W. Papageorge, Kevin Thom\
  *Journal of Political Economy*, 2020, vol. 128, No.4
author: "伊藤成朗"
engine: knitr
execute:
  echo: false
  freeze: auto
format:
  revealjs:
    standalone: true
    embed-resources: true
    slide-number: true
    incremental: true
    width: 1600
    height: 900
    preview-links: auto
    logo: GrootbergGiraffeHead.jpg
    footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/BPT.html>
    theme: [sky]
    linkcolor: red
    code-line-numbers: true
    code-copy: true
    code-block-border-left: "#31BAE9"
    toc: true
    toc-depth: 1
    template-partials: 
      - ../toc-slide.html
    includes:
      in_header: ../sky_add_FilePointer.html
      after-body: ../toc-add.html
    toc-bg-image: "../../GrootbergGiraffe3.jpg"
lightbox: auto
filters:
  - parse-latex
include-before-body:
#### latex math shorthand
  file: ../MathShorthand.html 
include-in-header:
  - text: |
      <style>
       .title {
        font-size: 2.0em;
      }
      </style>
#### From https://github.com/quarto-dev/quarto-cli/discussions/2951
#### I am happy with the left-indenting of the section line
#### Horizontal centering can be challenging (!) in quarto: stupid
#### https://github.com/quarto-dev/quarto-cli/issues/1231
#### It turns out an easy solution is :::{style="text-align: center;"}
  - text: |
      <style>
      .center-xy {
        margin: 0;
        position: absolute;
        top: 50%;
        left: 20%;
        -ms-transform: translateY(-50%), translateX(-50%);
        transform: translateY(-50%), translateX(-50%);
      }
      </style>
bibliography: ../seiro.bib
#### ../../../../../seiro.bib
#### copy c:/seiro/docs/seiro.bib ../seiro.bib
resources:
  - BPT.pdf
#### quarto render BPT.qmd
---
```{css, echo = F, results = "hide"}
Red {
  color: red;
}
Orange {
  color: #e94e2c;
}
Blue {
  color: blue;
}
.reveal .slides pre code {
    max-height: 800px;
}
#TOC {
  font-color: #e94e2c;
}
.big-code{
  font-size: 110%  
}
.divCenterCol
{
  text-align:center;
}
#### Below is suggested by Claude
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list ol {
####   list-style-type: none;
####   padding-left: 0;
####   margin-left: 0;
####   counter-reset: item;
#### }
#### .small-font-list ol li {
####   margin-bottom: 0.5em;
####   display: flex;
#### }
#### .small-font-list ol li::before {
####   content: counter(item) ". ";
####   counter-increment: item;
####   margin-right: 0.5em;
####   flex-shrink: 0;
#### }
#### .small-font-list ol li > p:first-child {
####   margin: 0;
####   display: inline;
#### }
.small-font-list {
  font-size: 70%;
  line-height: 1.2;
}
.small-font-list ol {
  list-style-type: none;
  padding-left: 0;
  margin-left: 0;
  counter-reset: item;
}
.small-font-list ol li {
  counter-increment: item;
  display: flex;
  align-items: flex-start;
  margin-bottom: 0.5em;
}
.small-font-list ol li::before {
  content: counter(item) ". ";
  flex: 0 0 2.5em;  /* Adjust this value to change the width of the number area */
  text-align: right;
  margin-right: 0.5em;
}
#### description list by Claude
/* Main container for the description list */
.description-list {
  font-size: 50% !important;  /* Reduces font size to 70% of the parent element */
  line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Styles for the description list element */
.description-list dl {
  margin-left: 0;  /* Removes default left margin */
  padding-left: 0;  /* Removes default left padding */
}
/* Styles for the term (header) in the description list */
.description-list dt {
  font-weight: bold;  /* Makes the term text bold */
  display: inline;  /* Displays the term inline. Uncomment for a line break after item header. */
  margin-right: 0.5em;  /* Adds some space after the term */
}
/* Styles for the description in the description list */
.description-list dd {
  display: inline;  /* Displays the description inline. Uncomment for a line break after item header. */
  margin-left: 0;  /* Removes default left margin */
}
/* Styles for paragraphs within the description */
.description-list dd p {
  display: inline;  /* Keeps paragraph content inline. Uncomment for a line break after item header. */
  margin: 0;  /* Removes default paragraph margins */
}
/* Creates a new line and adds space between list items */
.description-list dd::after {
  content: "";  /* Necessary for the pseudo-element to render */
  display: block;  /* Creates a line break */
  margin-bottom: 0.5em;  /* Adds space between items. Adjust as needed */
}
#### description with a line break between item header and texts
.description-lb {
  font-size: 70%;
  line-height: 1.2;
}
.description-lb dl {
  margin-left: 0;
  padding-left: 0;
}
.description-lb dt {
  font-weight: bold;
  margin-bottom: 0.2em;
}
.description-lb dd {
  margin-left: 0;
  margin-bottom: 1em;  /* Adjust this value to control spacing between items */
}
.description-lb dd p {
  margin: 0;
}
```

@BarthPapageorgeThom2020 take aways

* Gene-wealth gradient exists: EA score predicts wealth at retirement
* Gene-wealth gradient&lArr; education, income, bequests, parents’ education, but does not explain fully
* Gene-wealth gradient&rarr;0 when stock ownership is added
   * EA score $\SIM$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
* HHs without DB pension: gene-wealth gradient&uarr;

----

## SNP, polygenic score

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
Genome

![](DNAMacrostructure.jpg){height="100%"}

::: {style="font-size: 70%;"}
[https://courses.lumenlearning.com/...](https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication)
:::
::::

::::{.fragment}
![](DNAStructure.jpg){height="80%"}

::: {style="font-size: 70%;"}
_[https://www.haibunda.com/...](https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna-)_
:::
::::

:::::

----

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
![Double helix and rung](DoubleHelixRung.jpg)

::: {style="font-size: 70%;"}
[https://mammothmemory.net/...](https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html)
:::
::::

::::{.fragment}
![SNP](SNP.png)

::: {style="font-size: 70%;"}
_[https://studylib.net/...](https://studylib.net/doc/5786293/snp-presentation)_
:::
::::

:::::

----

<!-- tree.nathanfriend.io -->

```
.
└── Chromosome
    └── DNA
        └── Gene
            └── Genome
```

:::::{.columns}

::::{.column width="45%"}
![](genome.jpg)

::: {style="font-size: 70%;"}
[_Source: L. C. Douglas_](https://historiesofthingstocome.blogspot.com/2010/08/age-of-genome-4.html){height=5cm}
:::
::::

::::{.column width="45%"}
* Chromosomes: 1/2 each from both parents 
* DNA: Rolled up and shapes like an X
* Gene: DNA segments
* Genome: 3 billion nucletiodes (letters)
   * Letter sequence = genome sequence
   * 1.5 billion base pairs
::::

:::::

* Base pairs differ between people in less than 1% (15 million) locations  
* Single-nucleotide polymorphism (SNP) = such location
   * 2 chromosomes per person: AT-AT, GC-GC, AT-GC or 0, 2, 1 (reference = GC)

----

Twin studies

* Estimate how much genetic factors *collectively* matter for **explaining variations** of a trait  
   * Collectively: Unit = group level aggregation
   * Explaining variations: Variance decomposition 
* Do not reveal which SNP is correlated

. . .

Genome wide association studies (GWASs)

* Collect $J$ observable (?) SNPs of individual $i$.

. . .

Regress outcome $Y_{i}$ on each SNP using $J$ estimating equations:

$$
Y_{i}=\bfmu'\bfx_{i}+\beta_{j}SNP_{ij}+\epsilon_{ij}, \quad j=1,\dots,J.
$$

----

Polygenic score of $i$ for the outcome $Y$ = *"Educational attainment (EA) score"*

$$
PGS_{i}=\sum_{j=1}^{J}\tilde{\beta} SNP_{ij}
$$

* Use Bayesian LDpred procedure to correct for correlations in $\tilde{\beta}_{j}$

* Use all SNPs: Better out-of-sample results than using only SNPs with genome-wide significance $p$ value $< 5*10^{-8} =$ .0000005%

* PGS is considered to be a predictor of individual fixed effects
   * HRS sample: covariance with EA score / total schooling years variance = 10.6%
   * Edu attainment SNPs $\SIM$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
   * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]

----

## Interpretation

* Correlation, not causality: Outcomes &lArr; gene + environment (endogenous)
* Attenuation due to measurement errors in $\hat{\beta}_{j}$
* Linearity assumption &rarr; likely underestimation of gene contributions (relative to twin studies)
* Population stratification: Ethnic-group fixed effects &lArr; gene, environment  
   * Including first (10) principal components of SNP data as covariates is shown to control geographic variations
* External validity is up to HRS ancestral composition

----

## Data

1. Health and retirement survey: Age > 50 + partners, genetic samples 2006, 2008 
   * All households with at least one individual with genetic European 
   * "Retired households" in 1996, 1998, 2002, 2004, 2006, 2008, 2010
   * Households with 1-2 members, drop same sex HHs
   * Household-year observations of both members 65-75 years old
   * Self-reported earnings  
1. Social security administration data  
   * Income data, top coded

. . .

2590 HHs, 5701 HH-year observations (Table 1)

----

> our sample comprises households for whom wealth data are most likely to be both accurate and
comprehensive.

* All pension, annuity, social security income, defined-contribution retirement plans 
* Net value of housing, private businesses, vehicles, financial assets (cash, checking account, saving account, CD, stocks, mutual funds, trusts, other)
* Winsorize at 1st and 99th percentiles of log real total wealth

<br>

Standard controls
: age, sex of FR, calendar time, family member age-sex dummies and interactions

Full education controls
: education-age-sex dummies and interactions

----

## Selection biases

1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
   * Education $\SIM$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\SIM$ education gradient

----

## Selection biases

2. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
   * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\SIM$ education gradient &uarr; (Panel A vs. B)
   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\SIM$ education gradient &darr; (Panel A vs. C)
   * Smaller degree of selection leads to smaller q4-q1 differences
   * But larger measurement errors in wealth ($\because$ pensions, etc. are not measured)

----

> However, the magnitudes of these differences are similar and relatively modest across alternate samples. Restricting our sample to retired households balances concerns about sample selection and measurement error. 
[(p.19)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=19)

. . .

Meaning:

* We could have used a wider age band retired HH sample 
   * More precision (larger sample size, smaller measurement errors) 
   * Larger EA-edu gradient
* But we did not
   * Feel free to accept our results!

----

## Results

::::: {.columns}

::: {.column width="5%"}
:::{style="font-size: 45%;line-height: 1.1;"}
Fig 2A
:::
:::
::: {.column width="25%"}
![](Fig2A.jpg){width="70%"}
:::
::: {.column width="70%"}
EA score $\SIM$ wealth

*  EA score -1 to 1: $+$ $20,000

:::{.fragment}
Table 2 also shows EA score $\SIM$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
:::
:::

::: {.column width="5%"}
:::{.fragment}
:::{style="font-size: 45%;line-height: 1.1;"}
Fig 2B
:::
:::
:::
::: {.column width="25%"}
:::{.fragment}
![](Fig2B.jpg){width="70%"}
:::
:::

::: {.column width="70%"}
:::{.fragment}
EA score $\SIM$ wealth similar between high schoolers vs college grads, up to q3
:::
:::

::: {.column width="5%"}
:::{.fragment}
:::{style="font-size: 45%;line-height: 1.1;"}
Tab 4
:::
:::
:::
::: {.column width="25%"}
:::{.fragment}
![](Tab4.jpg){width="50%"}
:::
:::

::: {.column width="70%"}
:::{.fragment}
EA score $\SIM$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
:::

:::{.fragment}
Gradient: "ability" + school quality &larr; income measure error
:::
:::

:::::

----

Robust to:

::::: {.columns}

::: {.column width="50%"}
* $+$Non financial respondent EA scores
   * NFR EA scores is not partial-correlated
* $+$SSA income (more objective)
* $+$Non-retired HHs  
* $+$Self-reported income of HRS
:::

::: {.column width="40%"}
<div class="divCenterCol">
![](Tab5.jpg){width="50%"}
</div>
:::
:::::

* Changes in sample and definitions: HRS sampling weights, only 1 HH-year per sample, only coupled HHs, different age, complicated function of incomes, excluding retirement and housing wealth, different version of EA score 
* Adding: cognitive ability, number of children, death of HH member, years since retirement

----

## Gene-environment correlations

::::: {.columns}

::: {.column width="50%"}
Proxy of environment

* Direct: Bequests
* Indirect: Parental education

::: {.fragment}
Does not change the gene-wealth gradient
:::

* Parental investments are captured by respondent's schooling and labor incomes
:::

::: {.column width="40%"}
<div class="divCenterCol">
![](Tab6.jpg){width="50%"}
</div>
:::
:::::

----

## Additional mechanisms

EA score&uarr;

::::: {.columns}

::: {.column width="60%"}
* One-year mortality&darr; for females 
* Expected longeviity&uarr; for females 
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](Tab7.jpg){width="50%"}
</div>
:::

::: {.column width="60%"}
* Elicited risk aversion&darr; in gambles on incomes, business, but not on inheritance 
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](Tab8.jpg){width="50%"}
</div>
:::

::: {.column width="60%"}
* Stock ownership&uarr; but not of other assets 
* EA score $\SIM$ wealth gradient &rarr;0 if we add stock ownership  
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](Tab9.jpg){width="50%"}
</div>
:::

:::::

----

EA score&uarr;

::::: {.columns}

::: {.column width="60%"}
* Deviation from objective probability&darr; 
* Indication of complete certainty&darr;  
* Financial planning time horizon&uarr; (viz. a few months) (except longer than 10 years)
   * Better able to think about complex and abstract decisions
   * @GabaixLaibson2022: Precise signals &rArr; behaves as if more patient
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](Tab10.jpg){width="90%"}
</div>
:::

::: {.column width="60%"}
* Elicited stock returns&uarr; &rArr; stock ownership&uarr;, wealth&uarr; 
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](TabS9.jpg){width="50%"}
</div>
:::

::: {.column width="60%"}
* Robust to adding cognitive test scores 
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](TabS15.jpg){width="40%"}
</div>
:::

:::::

----

::::: {.columns}

::: {.column width="60%"}
* EA score $\SIM$ wealth, conditional on DB pension participation
* EA score $\SIM$ wealth&uarr; if not participating in DB pension
:::

::: {.column width="30%"}
<div class="divCenterCol">
![](Tab11.jpg){width="90%"}
</div>
:::

:::::

:::{.fragment}
Gene-wealth gradient &lArr; a household's <Red>difficulty in making wise financial decisions</Red>
:::

----

## Conclusions

* Previous studies: EA scores $\SIM$ education
* This study: EA scores $\SIM$ wealth
* EA scores $\SIM$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
   * Teaching probability may make human capital more homogenous?
   * Not necessarily, gene-wealth gradient is a biased estimate &larr; environment
   * But it hints something can be done there

. . .

::: {style="font-size: 80%;"}
   > For example, it may be that children with lower polygenic scores begin to face challenges at particular ages or struggle to meet specific educational milestones. In that case,we could better target educational policies to help alleviate these road-blocks. 
:::

こうした学習の遅れが格差の源泉と認識できるようになった

----

## Conclusions

* "Ignoring heterogeneity in EA scores may place too much weight on education and inheritance"
   * Using HH (parent-multiple children) fixed effect estimator should prevent this? 
* Intervention changes the nature contribution (gradient size): DB pensions

----

## 感想

* 「貧しさは遺伝する」?...PGSを使って資産との相関を計測した研究
* 強い根拠はないにせよ、介入ポイントを議論したのは素晴らしい
* Gene-wealth gradient = corr[PGS, wealth|(endogenous) variables]  
* 偏相関が小さくなるポイント: 株式所有 &lArr; longer time horizon, optimism
* 偏相関が大きくなるポイント: 確定給付年金
    * 株式所有と確定給付年金を加えたら偏相関は?
* 金融教育? 
* 確定給付年金をもっと広く手厚く(できるのか?)

----