### 2025-01-07 posts/RustichiniIaconoLeeMcGue/RILM_PGSEdu_JPE2023.qmd
initial commit

---
title: "Educational attainment and intergenerational mobility: A polygenic score analysis"
subtitle: Aldo Rustichini, William G. Iacono, James J. Lee, and Matt McGue|
  Journal of Political Economy 2023 131:10, 2724-2779
author: "伊藤成朗"
engine: knitr
execute:
  echo: false
  freeze: auto
format:
  revealjs:
    standalone: true
    embed-resources: true
    slide-number: true
    incremental: true
    width: 1600
    height: 900
    preview-links: auto
    logo: GrootbergGiraffeHead.jpg
    footer: <http://seiroito.github.io/RILM_PGSEdu_JPE2023.html>
    css: ../../../../seiro.css
    #linkcolor: red#marked off because sky-add sets it to DarkGreen
    theme: [sky, sky-add]
    code-line-numbers: true
    code-copy: true
    code-block-border-left: "#31BAE9"
    toc: true
    toc-depth: 1
    template-partials: 
      - ../../../../toc-slide.html
    include-after-body: ../../../../toc-add.html
    toc-bg-image: "../../GrootbergGiraffe3.jpg"
lightbox: 
  match: auto
  desc-position: bottom
filters:
  - parse-latex
include-in-header:
  - text: |
      <style>
       .title {
        font-size: 2.0em;
      }
      </style>
#### From https://github.com/quarto-dev/quarto-cli/discussions/2951
#### I am happy with the left-indenting of the section line
#### Horizontal centering can be challenging (!) in quarto: stupid
#### https://github.com/quarto-dev/quarto-cli/issues/1231
#### It turns out an easy solution is :::{style="text-align: center;"}
  - text: |
      <style>
      .center-xy {
        margin: 0;
        position: absolute;
        top: 50%;
        left: 20%;
        -ms-transform: translateY(-50%), translateX(-50%);
        transform: translateY(-50%), translateX(-50%);
      }
      </style>
include-before-body:
#### latex math shorthand
  file: ../../../../MathShorthand.html 
bibliography: c:/seiro/docs/seiro.bib
#### ../../../../seiro.bib
resources:
  - RILM_PGSEdu_JPE2023.pdf
# quarto render RILM_PGSEdu_JPE2023.qmd
---
```{css, echo = F, results = "hide"}
TallerOutput {
  height: 800px;
}
.small-font-list {
  font-size: 70%;
  line-height: 1.2;
}
.small-font-list ol {
  list-style-type: none;
  padding-left: 0;
  margin-left: 0;
  counter-reset: item;
}
.small-font-list ol li {
  counter-increment: item;
  display: flex;
  align-items: flex-start;
  margin-bottom: 0.5em;
}
.small-font-list ol li::before {
  content: counter(item) ". ";
  flex: 0 0 2.5em;  /* Adjust this value to change the width of the number area */
  text-align: right;
  margin-right: 0.5em;
}
#### description list by Claude
/* Main container for the description list */
.description-list {
  font-size: 50% !important;  /* Reduces font size to 70% of the parent element */
  line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Styles for the description list element */
.description-list dl {
  margin-left: 0;  /* Removes default left margin */
  padding-left: 0;  /* Removes default left padding */
}
/* Styles for the term (header) in the description list */
.description-list dt {
  font-weight: bold;  /* Makes the term text bold */
  display: inline;  /* Displays the term inline. Uncomment for a line break after item header. */
  margin-right: 0.5em;  /* Adds some space after the term */
}
/* Styles for the description in the description list */
.description-list dd {
  display: inline;  /* Displays the description inline. Uncomment for a line break after item header. */
  margin-left: 0;  /* Removes default left margin */
}
/* Styles for paragraphs within the description */
.description-list dd p {
  display: inline;  /* Keeps paragraph content inline. Uncomment for a line break after item header. */
  margin: 0;  /* Removes default paragraph margins */
}
/* Creates a new line and adds space between list items */
.description-list dd::after {
  content: "";  /* Necessary for the pseudo-element to render */
  display: block;  /* Creates a line break */
  margin-bottom: 0.5em;  /* Adds space between items. Adjust as needed */
}

#### description with a font size control (not effective)
/* Main container for the description list */
.des-list-boxed {
    font-size: 60% !important;  /* Default font size */
    line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Specific font size override */
.des-list-boxed[data-font-size] {
    font-size: attr(data-font-size string, 60%) !important;
}
.des-list-boxed dl {
    margin-left: 0;  /* Removes default left margin */
    padding-left: 0;  /* Removes default left padding */
    width: 100%;  /* Ensures the container takes full width */
}
.des-list-boxed dt {
    font-weight: bold;  /* Makes the term text bold */
    text-align: right;  /* Right aligns the term text */
    padding-right: 1em;  /* Adds spacing after the term */
    float: left;  /* Allows description to wrap around header */
    white-space: nowrap;  /* Prevents item header from wrapping */
}
.des-list-boxed dd {
    margin-left: 0;  /* Removes default left margin */
    overflow: hidden;  /* Clears float */
}
.des-list-boxed dd p {
    margin: 0;  /* Removes default paragraph margins */
    word-wrap: break-word;  /* Allows description to wrap */
    overflow-wrap: break-word;  /* Modern alternative to word-wrap */
}
.des-list-boxed dd p:first-child {
    /* First line of description */
    /* No special styling needed, will naturally wrap around header */
}
.des-list-boxed dd p:not(:first-child) {
    /* Lines after the first line */
    display: block;  /* Ensures full width */
    margin-left: 1em;  /* Indentation for subsequent lines */
}
.des-list-boxed dd::after {
    content: "";  /* Necessary for the pseudo-element to render */
    display: block;  /* Creates a line break */
    margin-bottom: 0.5em;  /* Adds space between items */
    clear: both;  /* Clears the float */
}

#### description with a line break between item header and texts
.description-lb {
  font-size: 70%;
  line-height: 1.2;
}
.description-lb dl {
  margin-left: 0;
  padding-left: 0;
}
.description-lb dt {
  font-weight: bold;
  margin-bottom: 0.2em;
}
.description-lb dd {
  margin-left: 0;
  margin-bottom: 1em;  /* Adjust this value to control spacing between items */
}
.description-lb dd p {
  margin: 0;
}
gray {
  color: #aeb6bf;
}
```

## Summary of [this paper](https://doi.org/10.1086/724860)

* US Minnesota Twin Study Sample
* Build a genetic model of human capital investments with genotype transmission (with better genetics foundation than *ad hoc* standard econ models)
* Intergenerational elaciticity of income is predicted to be larger in genetic models
* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect on child human capital
* Child PGS is relevant
* Genetic nurture (introverted genotype produces an environment to grow introverted child, independently of income and education) is not found

----

:::{style="text-align: center;"}
# Background
:::

----

----

## SNP, polygenic score

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
Genome

![](DNAMacrostructure.jpg){height="100%"}

::: {style="font-size: 70%;"}
[https://courses.lumenlearning.com/...](https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication)
:::
::::

::::{.fragment}
![](DNAStructure.jpg){height="80%"}

::: {style="font-size: 70%;"}
_[https://www.haibunda.com/...](https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna-)_
:::
::::

:::::

----

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
![Double helix and rung](DoubleHelixRung.jpg)

::: {style="font-size: 70%;"}
[https://mammothmemory.net/...](https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html)
:::
::::

::::{.fragment}
![SNP](SNP.png)

::: {style="font-size: 70%;"}
_[https://studylib.net/...](https://studylib.net/doc/5786293/snp-presentation)_
:::
::::

:::::

----

<!-- tree.nathanfriend.io -->

```
.
└── Chromosome
    └── DNA
        └── Gene
            └── Genome
```

:::::{.columns}

::::{.column width="40%"}
![](genome.jpg)

::: {style="font-size: 70%;"}
[_Source: L. C. Douglas_](https://historiesofthingstocome.blogspot.com/2010/08/age-of-genome-4.html){height=5cm}
:::
::::

::::{.column width="60%"}
* Chromosomes: 1/2 each from both parents 
* DNA: Rolled up and shapes like an X
* Gene: DNA segments
* Genome: 3 billion nucletiodes (letters)
   * Letter sequence = genome sequence
   * 1.5 billion base pairs (alelles, 対立遺伝子)
::::

:::::

* Base pairs differ between people in less than 1% (15 million) locations  
* Single-nucleotide polymorphism (SNP) = such location
   * 2 chromosomes per person &rarr; AT-AT, GC-GC, AT-GC or 0, 2, 1 (reference = GC)

----

Twin studies

* Estimate how much genetic factors *collectively* matter for **explaining variations** of a trait  
   * Collectively: Unit = group level aggregation
   * Explaining variations: Variance decomposition 
* Do not reveal which SNP is correlated

----

Genome wide association studies (GWASs)

* Collect $M$ observable (?) SNPs of (family $i$) individual $j$.

. . .

Regress outcome $y^{i}_{j}$ on $m$-th SNP for all $m=1, \dots, M$:

$$
y^{i}_{j}
=
\beta_{m}SNP^{i}_{jm}+\epsilon^{i}_{jm}, \quad m=1,\dots,M.
$$

Remember:

* Environment is endogenous to the trait in SNP regressions
   * $\cov[SNP^{i}_{jm}, \epsilon^{i}_{jm}]> 0$: Parents see traits which may be correlated with $SNP^{i}_{jm}$ and decide on investments 
   * Estimated $\beta_{m}$ should be upwardly biased
* No attempts to correct for this

----

Polygenic score = Predicted outcomes based on all relevant SNPs

$$
PGS^{i}_{j}=\sum_{m=1}^{M}\tilde{\beta}_{m} SNP^{i}_{jm}
$$

* Use Bayesian LDpred procedure to correct for correlations in $\tilde{\beta}_{m}$
  * @BarthPapageorgeThom2020 follow the convention and use all SNPs: Better out-of-sample results than using only SNPs with genome-wide significance $p$ value $< 5*10^{-8} =$ .0000005%

* PGS is considered to be a predictor of individual fixed effects
   * Edu attainment SNPs $\SIM$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
   * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]
   * HRS sample: cov[EA score, schooling years] / variance[schooling years] = 10.6%

----

## Intergenerational skill formation

* @BeckerTomes1979: AR(1)
* @Goldberger1989: Weighted average of all ancestors
* Both are incorrect, this paper shows there exists an invariant (stable) mapping father genotype distribution × mother genotype distribution $\to$ child genotype distribution, so skill formation depends on parental genotype distributions

----

```{mermaid}
%%| label: fig-mermaid
%%| fig-width: 6
%%| fig-cap: |
%%|   How parents affect child education outcomes
flowchart LR
  A1[father] --> A3(couple)
  A2[mother] --> A3(couple)
  A3[couple] --> B("child
   genotype")
  A3[couple] --"genetic nurture"--> C("child
   environment")
  B("child
   genotype") --> D("child
    education")
  C("child
   environment") <--"gene×environ"--> B("child
    genotype")
  C("child
   environment") --> D("child
    education")
```

----

:::{style="text-align: center;"}
# Model 
:::

----

::: {style="font-size: 80%;"}
Family $i$ parents' problem with twins $j=1, 2$ of observable skills $\theta^{i}_{j}$
$$
\begin{aligned}
\max_{\{E^{i}, I^{i}_{1}, I^{i}_{2}\}} \;\;\;&\E\left[(1-\delta)\ln E^{i}+\delta\left(y^{i}_{1}+y^{i}_{2}\right)\left|\theta^{i}_{1}, \theta^{i}_{2}\right.\right]\\
\st \;\;\; & E^{i}+I^{i}_{1}+I^{i}_{2}=e^{y^{i}}
\end{aligned}
\tag{6}
$$
$$
\begin{align}
\scriptsize{\mbox{Human capital production function}} && 
h^{i}_{j}&=\alpha_{I}\ln I^{i}_{j}+\alpha_{\theta}\theta^{i}_{j}+\epsilon^{\mathrm{h}i}_{j} && \phantom{mmmmmmmmmmmm} \tag{8}\label{eq8}\\
\scriptsize{\mbox{Child income}} && 
y^{i}_{j}&=\alpha_{\mathrm{h}}h^{i}_{j}+\epsilon^{\mathrm{y}i}_{j} &&\tag{9}\label{eq9}
\end{align}
$$
Solution
$$
I^{i}=\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}e^{y^{i}}
\tag{11}
$$
Optimised child income
$$
y^{i}_{j}=\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\ln 
\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}
+\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta^{i}_{j}+\alpha_{\mathrm{h}}\epsilon^{\mathrm{h}i}_{j}+\epsilon^{\mathrm{y}i}_{j}
\tag{12}\label{eq12}
$$ 
:::

----

Matching of parents

* Assume independence of genotype conditional on observable characteristics [incomes, skills $w(g)$]
   * Matches are made on observables, not on genotype
   * Matches are random within obserbable strata *"worth" classes*=$w_{\mathrm{y}}y+w(g)$]
      * Genotype matching is random within a stratum
   * Matches are nonrandom at the population due to population stratification based on observables

. . .

Genetic transmission: $\Delta$(father genotype) × $\Delta$(mother genotype) $\to$ $\Delta$(child genotype)

The appendix D of paper shows an invariant mapping exists with properties shown in Theorem 3.1

----

Skills production function (genotype enters here, hence the parental genotypes)
$$
\theta^{i}_{j}=w\left(g^{i}_{j}\right)+\pi y^{i}+\bfPi\bfx^{i}_{j}+F^{i}+\epsilon^{\theta i}_{j}
\tag{4}\label{eq4}
$$
Estimating equations: Substitute \eqref{4} into  \eqref{12} gives child income decomposition, and \eqref{11} into  \eqref{8} gives human capital decomposition
$$
\begin{align}
y^{i}_{j}
&=
\alpha_{\theta}\alpha_{\mathrm{h}}w\left(g^{i}_{j}\right)
+\left(\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}+\alpha_{\theta}\alpha_{\mathrm{h}}\pi\right)y^{i}
+\alpha_{\theta}\alpha_{\mathrm{h}}F^{i}\\
&\hspace{1em}
+\alpha_{\theta}\alpha_{\mathrm{h}}\bfPi\bfx^{i}_{j}
+\alpha_{\theta}\alpha_{\mathrm{h}}\epsilon^{\theta i}_{j}
+\alpha_{\mathrm{h}}\epsilon^{\mathrm{h} i}_{j}
+\epsilon^{\mathrm{y}i}_{j}\tag{21}\label{eq21}\\
h^{i}_{j}
&=
\alpha_{\theta}w\left(g^{i}_{j}\right)
+\left(\alpha_{\mathrm{I}}+\alpha_{\theta}\pi\right)y^{i}
+\alpha_{\theta}F^{i}\\
&\hspace{1em}
+\alpha_{\theta}\bfPi\bfx^{i}_{j}
+\alpha_{\theta}\epsilon^{\theta i}_{j}
+\epsilon^{\mathrm{y}i}_{j}\tag{23}\label{eq23}
\end{align}
$$

----

Can show IGE is smaller in "standard model" than genetic model

::: {style="font-size: 80%;"}
@BeckerTomes1979
$$
\begin{align}
\theta_{t+1}
&=
\eta \theta_{t}+\epsilon^{\theta}_{t+1}\tag{26}\label{eq26}\\
y^{i}_{t+1}
&=
\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}_{t}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{25}\label{eq25}
\end{align}
$$
Genetic model of this paper assumes
$$
\begin{align}
\theta
&=
w^{\theta}_{\mathrm{f}}\theta_{\mathrm{f}}+w^{\theta}_{\mathrm{m}}\theta_{\mathrm{m}}\tag{A4}\label{eqA4}\\
y^{i}
&=
w^{\mathrm{y}}_{\mathrm{f}}y^{i}_{\mathrm{f}}+w^{\mathrm{y}}_{\mathrm{m}}y^{i}_{\mathrm{m}}\tag{A3}\label{eqA3}
\end{align}
$$
Plugging these give
$$
\begin{align}
\theta_{t+1}
&=
\eta \sum_{s=m,f}w^{\theta}_{s}\theta_{st}+\epsilon^{\theta}_{t+1}\tag{33}\label{eq33}\\
y^{i}_{t+1}
&=
\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\sum_{s=m,f}w^{\mathrm{y}}_{s}y^{i}_{st}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{32}\label{eq32}
\end{align}
$$
:::

----

:::{style="text-align: center;"}
# Estimation strategy 
:::

----

Structural equation model (SEM)

* Structural equations
* Measurement (error) equations

. . .

Identification conditions of structural parameters remain the same as in simultaneous equation models

In this paper:

* identification for only recursive model is discussed (Tab 2)
* use of IVs is not discussed (Tab 3)

. . .

Controls used in estimation

* 10 principal components of PGS
* Parent-child age difference

----

Identification ideas

1. Dizygotic twins [Table 1, 2]
   * Share birth days, environment
   * Differ in genotypes
   * Ideal to see impacts of genotypes on outcomes

----

Identification ideas

2. SEM
   * Recursive system (37)-(42) are identified as usual [Table 2]
   * Structural parameters of endogenous variables are recovered from \eqref{eq43}-\eqref{eq45} (how?) [Table 3]
      * Given PGS of child, PGS of parents should be redundant, so long as they do not affect environment (gene-environment correlation, or "passive rGE")

. . .

::: {style="font-size: 80%;"}
$$
\begin{align}
e^{i}_{\mathrm{h}}
&=
\alpha_{e_{\mathrm{h}}}+
\gamma^{\mathrm{pGS}_{m}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{e_{\mathrm{h}}}\tag{43}\label{eq43}\\
y^{i}_{\mathrm{h}}
&=
\alpha_{y_{\mathrm{h}}}+\gamma^{\mathrm{pGS}_{m}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{y_{\mathrm{h}}}\tag{44}\label{eq44}\\
e^{i}_{j}
&=
\alpha_{e}+\gamma^{e_{\mathrm{h}}}_{\mathrm{e}}e_{\mathrm{h}}^{i}
+\gamma^{y_{\mathrm{h}}}_{\mathrm{e}}y_{\mathrm{h}}^{i}
+\gamma^{\mathrm{pGS}}_{\mathrm{e}}\mathrm{pGS}^{i}_{j}
+\gamma^{\mathrm{pGS}_{m}}_{\mathrm{e}}\mathrm{pGS}^{i}_{m}
+\gamma^{\mathrm{pGS}_{f}}_{\mathrm{e}}\mathrm{pGS}^{i}_{f}
+\xi_{\mathrm{e}}\tag{45}\label{eq45}
\end{align}
$$
:::

----

::: {style="font-size: 70%;"}
**Data:** Minnesota Twin Study

* Cognitive measures to estimate IQ
   * Wechsler Intelligence Scale for Children-Revised (WISC-R)
   * Wechsler Adult Intelligence Scale-Revised (WAIS-R)
* Noncognitive measures 
   * MPQ (Mulitdimensional Personality Questionnaire)
      * PA (positive emotionality of affectivity; well-being, social potency, achievement, social closeness)
      * NA (negative emotionality of affectivity; stress reaction, alienation, aggression)
      * constraint (control, harm avoidance, traditionalism)
   * DSM-IV (Diagnostic and Statisitcal Manual of Mental Disorders, fourth edition)
      * Externalizing: defiant disorder, conduct disorder, adult antisocial behavior
      * Academic effort: 8 items answered by twin mothers
      * Academic problems: 3 items answered by twin mothers
* Family background
   * Hollingsworth scale on occupational status
   * Four year college degrees of parents
:::

----

:::{style="text-align: center;"}
# Estimation results 
:::

----

:::::::::::{.columns}

::::{.column width="45%"}
![IGE](Tab1.jpg)
::::

::::{.column width="55%"}
* IGE = .13 is small relative to previous studies (.2 ~ .3)
   * Attenuation (measurement errors in family incomes)
   * European population has lower IGEs, Sweden = .125
   * Reduced by `r round((1-(.78/.134))*100, 0)`% when PGS and mediators are added
* Effects of PGS
   * = .078 in (2)
   * = 0 in (3) when mediators of PGS are added
::::

::::::::::

----

:::::::::::{.columns}

::::{.column width="45%"}
![recursive SEM](Tab2.jpg)
::::

::::{.column width="55%"}
Education years&uarr; with

* Cognitive ability, noncognitive ability
   * PGS of child, parents are not related when parental education, family income are added
* Parental income, education
   * Impacts are twice as large for parental education

Cognitive ability&uarr; with

* PGS

Noncognitive ability&uarr; with

* PGS, smaller magnitude
::::

::::::::::

----

----

:::::::::::{.columns}

::::{.column width="45%"}
![full SEM](Tab3.jpg)
::::

::::{.column width="55%"}
Dizygotic (fraternal) + monozygotic (identical) twins

Education years&uarr; with

* PGS of child and mother
   * PGS of father is unrelated
* Parental income, education
   * Magnitude is larger for parental education

Parental education years&uarr; with

* PGS of parents
   * Magnitude is larger for father PGS

Family income&uarr; with

* PGS of parents
   * Magnitude is larger for father PGS
::::

::::::::::

----

----

:::::::::::{.columns}

::::{.column width="25%"}
![](TabS6.jpg){width="50%"}

![](TabS7.jpg){width="50%"}

![](TabS8.jpg){width="50%"}

![](TabS9.jpg){width="50%"}
::::

::::{.column width="75%"}
* PGS of parents are not related to outcomes once IQ, soft skills index, parental education and family income are added
* Most parental impacts on education are through incomes and parental education
* No other indirect pathway remain from parental PGS, after parental education and incomes are addressed
   *  No passive rGE other than parental education and incomes (genetic nurture) is found
::::

::::::::::

----

感想

* 咀嚼できていません
* 人的資本投資を内生化したsteady state genetic transmission mappingの存在を証明
* それと実証部分の結びつきが弱い(理解できていない)
* 親の教育と所得を考慮すると、遺伝情報は子の教育に影響しないことを実証的に示したのは斬新
* SEMはidentification assumptionが不明(recursiveモデル以外)、IVを使っていない?
* Multiple testing気にせず

----

#### References

::: {#refs}
:::