### 2025-06-18 posts/DelabastitaRubens/DelabastitaRubens2.qmd

---
title: "Delabastita and Rubens 2025"
author: "claude 4 sonnet, instructed by me"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
####  pdf: default
  docx: default
  epub: default
theme: 
  light: sky
  dark: darkly
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render DelabastitaRubens2.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/DelabastitaRubens/'; rmarkdown::render(paste0(path, 'DelabastitaRubens.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  ####color: #010101;/*文字色*/
  color: #87a5f2;
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

# Equations and Variable Definitions from Coal Mining Study

## Main Equations

### Equation (1) - Cobb-Douglas Production Function
$$
q_{ft} = \beta_{l} l_{ft} + \beta_{m} m_{ft} + \beta_{k} k_{ft} + \omega_{ft} \tag{1}
$$

### Equation (2) - TFP Transition Process
$$
\omega_{ft} = \rho \omega_{ft-1} + u_{ft} \tag{2}
$$

### Labor Supply Function - Equation (3)
$$
W^{l}_{it} = L_{it}^{\Psi^{l}} \eta_{it} \tag{3}
$$

### Cost Minimization Objective - Equation (4)
$$
\min_{L_{ft}, M_{ft}} \left\{ \sum_{g \in F_{i(f)t}} \left( \lambda_{fgt} (L_{gt} W^{l}_{gt} + M_{gt} W^{m}_{gt}) \right) - MC_{ft} \left( Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) - Q_{ft} \right) \right\} \tag{4}
$$

### No Collusion Cost Minimization - Equation (5)
$$
\min_{L_{ft}, M_{ft}} \left( L_{ft} W^{l}_{ft} + M_{ft} W^{m}_{ft} \right) - MC_{ft} \left( Q_{ft} - Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) \right) \tag{5}
$$

### First-Order Condition for Labor (No Collusion) - Equation (6)
$$
L_{ft} \frac{\partial W^{l}_{it}}{\partial L_{it}} + W^{l}_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{6}
$$

### Markdown Under No Collusion - Equation (7)
$$
\widetilde{m}^{l}_{ft} = 1 + s^{l}_{ft} \Psi^{l} \tag{7}
$$

### Perfect Collusion Cost Minimization - Equation (8)
$$
\min_{L_{ft}, M_{ft}} \left\{ \sum_{g \in F_{i(f)t}} \left( L_{gt} W^{l}_{gt} + M_{gt} W^{m}_{gt} \right) - MC_{ft} \left( Q_{ft} - Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) \right) \right\} \tag{8}
$$

### First-Order Condition for Labor (Perfect Collusion) - Equation (9)
$$
L_{it} \frac{\partial W^{l}_{it}}{\partial L_{it}} + W^{l}_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{9}
$$

### Markdown Under Perfect Collusion - Equation (10)
$$
\overline{m}^{l}_{ft} = 1 + \Psi^{l} \tag{10}
$$

### General First-Order Condition - Equation (11)
$$
W^{l}_{it} + \widetilde{\lambda}_{ft} \frac{\partial W^{l}_{it}}{\partial L_{it}} L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{11}
$$

### General Markdown Expression - Equation (12)
$$
m^{l}_{ft} = 1 + \widetilde{\lambda}_{ft} \Psi^{l} \tag{12}
$$

### Markup from Materials - Equation (13)
$$
\mu_{ft} = \frac{\beta_{m}}{\alpha^{m}_{ft}} \tag{13}
$$

### Core Identification Equation - Equation (14)
$$
m^{l}_{ft} = 1 + \widetilde{\lambda}_{ft} \Psi^{l} = \frac{\beta_{l} \alpha^{m}_{ft}}{\beta_{m} \alpha^{l}_{ft}} \tag{14}
$$

### Conduct Parameter Definition - Equation (15)
$$
\widehat{\lambda}_{ft} = \frac{m^{l}_{ft} - \widetilde{m}^{l}_{ft}}{\overline{m}^{l}_{ft} - \widetilde{m}^{l}_{ft}} \tag{15}
$$

### Moment Conditions for Production Function - Equation (16)
$$
E[u_{ft} | (l_{fr-1}, m_{fr-1}, k_{fr}, w^{agri}_{r-1})_{r \in \{2,...,t\}}] = 0 \tag{16}
$$

### GMM Moment Conditions - Equation (17)
$$
E[q_{ft} - \rho q_{ft-1} - \beta_{0}(1-\rho) - \beta_{l}(l_{ft} - \rho l_{ft-1}) - \beta_{m}(m_{ft} - \rho m_{ft-1}) - \beta_{k}(k_{ft} - \rho k_{ft-1}) | (l_{ft-1}, m_{ft-1}, k_{ft}, k_{ft-1}, w^{agri}_{t-1})] = 0 \tag{17}
$$

## Additional Key Definitions

### Markdown Definition
$$
m^{l}_{ft} \equiv \frac{MRPL_{ft}}{W^{l}_{ft}}
$$

### Percentage Markdown Definition
$$
d^{l}_{ft} \equiv \frac{MRPL_{ft} - W^{l}_{ft}}{MRPL_{ft}} = \frac{m^{l}_{ft} - 1}{m^{l}_{ft}}
$$

### Markup Definition
$$
\mu_{ft} \equiv \frac{P_{ft}}{MC_{ft}}
$$

### Labor Market Share
$$
s^{l}_{ft} \equiv \frac{L_{ft}}{L_{it}}
$$

### Revenue Share of Labor
$$
\alpha^{l}_{ft} \equiv \frac{W^{l}_{ft} L_{ft}}{P_{ft} Q_{ft}}
$$

### Revenue Share of Materials
$$
\alpha^{m}_{ft} \equiv \frac{W^{m}_{ft} M_{ft}}{P_{ft} Q_{ft}}
$$

### Inverse Labor Supply Elasticity (Firm-level)
$$
\theta^{l}_{ft} \equiv \frac{\partial W^{l}_{ft}}{\partial L_{ft}} \frac{L_{ft}}{W^{l}_{ft}}
$$

### Inverse Materials Supply Elasticity (Firm-level)
$$
\theta^{m}_{ft} \equiv \frac{\partial W^{m}_{ft}}{\partial M_{ft}} \frac{M_{ft}}{W^{m}_{ft}}
$$

## Variable Definitions

| Variable | Definition |
|----------|------------|
| $Q_{ft}$ | Output (tonnage of coal) extracted by firm $f$ in year $t$ |
| $L_{ft}$ | Amount of effective labor throughout the year for firm $f$ in year $t$ |
| $M_{ft}$ | Amount of intermediate inputs purchased by firm $f$ in year $t$ |
| $K_{ft}$ | Capital stock (steam engines) used by firm $f$ in year $t$ |
| $q_{ft}$ | Logarithm of output $Q_{ft}$ |
| $l_{ft}$ | Logarithm of labor $L_{ft}$ |
| $m_{ft}$ | Logarithm of materials $M_{ft}$ |
| $k_{ft}$ | Logarithm of capital $K_{ft}$ |
| $\beta_{l}$ | Output elasticity of labor |
| $\beta_{m}$ | Output elasticity of materials |
| $\beta_{k}$ | Output elasticity of capital |
| $\omega_{ft}$ | Log total factor productivity |
| $u_{ft}$ | Unexpected productivity shock |
| $\rho$ | Serial correlation parameter in productivity process |
| $W^{l}_{it}$ | Wage in labor market $i$ in year $t$ |
| $W^{l}_{ft}$ | Wage for firm $f$ in year $t$ |
| $W^{m}_{ft}$ | Price of materials for firm $f$ in year $t$ |
| $L_{it}$ | Market-level employment in market $i$ in year $t$ |
| $\Psi^{l}$ | Inverse market-level labor supply elasticity |
| $\eta_{it}$ | Market-specific residual in labor supply |
| $\lambda_{fgt}$ | Collusion weight that firm $f$ puts on firm $g$'s costs |
| $\widetilde{\lambda}_{ft}$ | Conduct parameter (firm-level aggregate of bilateral conduct parameters) |
| $\widehat{\lambda}_{ft}$ | Normalized conduct parameter ranging from 0 to 1 |
| $MC_{ft}$ | Marginal cost for firm $f$ in year $t$ |
| $P_{ft}$ | Coal price for firm $f$ in year $t$ |
| $MRPL_{ft}$ | Marginal revenue product of labor for firm $f$ in year $t$ |
| $m^{l}_{ft}$ | Wage markdown (ratio of MRPL to wage) |
| $\widetilde{m}^{l}_{ft}$ | Wage markdown under no collusion |
| $\overline{m}^{l}_{ft}$ | Wage markdown under perfect collusion |
| $d^{l}_{ft}$ | Percentage wage markdown |
| $\mu_{ft}$ | Price markup (ratio of price to marginal cost) |
| $s^{l}_{ft}$ | Labor market share of firm $f$ |
| $\alpha^{l}_{ft}$ | Revenue share of labor for firm $f$ in year $t$ |
| $\alpha^{m}_{ft}$ | Revenue share of materials for firm $f$ in year $t$ |
| $\theta^{l}_{ft}$ | Inverse firm-level labor supply elasticity |
| $\theta^{m}_{ft}$ | Inverse firm-level materials supply elasticity |
| $F_{i(f)t}$ | Set of firms in market $i$ (where firm $f$ operates) in year $t$ |
| $w^{agri}_{t-1}$ | Agricultural wages in Belgium in year $t-1$ (instrument) |

## Tab-Separated Variable Definitions

```
Variable	Definition
Q_ft	Output (tonnage of coal) extracted by firm f in year t
L_ft	Amount of effective labor throughout the year for firm f in year t
M_ft	Amount of intermediate inputs purchased by firm f in year t
K_ft	Capital stock (steam engines) used by firm f in year t
q_ft	Logarithm of output Q_ft
l_ft	Logarithm of labor L_ft
m_ft	Logarithm of materials M_ft
k_ft	Logarithm of capital K_ft
beta_l	Output elasticity of labor
beta_m	Output elasticity of materials
beta_k	Output elasticity of capital
omega_ft	Log total factor productivity
u_ft	Unexpected productivity shock
rho	Serial correlation parameter in productivity process
W_l_it	Wage in labor market i in year t
W_l_ft	Wage for firm f in year t
W_m_ft	Price of materials for firm f in year t
L_it	Market-level employment in market i in year t
Psi_l	Inverse market-level labor supply elasticity
eta_it	Market-specific residual in labor supply
lambda_fgt	Collusion weight that firm f puts on firm g's costs
lambda_tilde_ft	Conduct parameter (firm-level aggregate of bilateral conduct parameters)
lambda_hat_ft	Normalized conduct parameter ranging from 0 to 1
MC_ft	Marginal cost for firm f in year t
P_ft	Coal price for firm f in year t
MRPL_ft	Marginal revenue product of labor for firm f in year t
m_l_ft	Wage markdown (ratio of MRPL to wage)
m_l_tilde_ft	Wage markdown under no collusion
m_l_bar_ft	Wage markdown under perfect collusion
d_l_ft	Percentage wage markdown
mu_ft	Price markup (ratio of price to marginal cost)
s_l_ft	Labor market share of firm f
alpha_l_ft	Revenue share of labor for firm f in year t
alpha_m_ft	Revenue share of materials for firm f in year t
theta_l_ft	Inverse firm-level labor supply elasticity
theta_m_ft	Inverse firm-level materials supply elasticity
F_i_f_t	Set of firms in market i (where firm f operates) in year t
w_agri_t_minus_1	Agricultural wages in Belgium in year t-1 (instrument)
```

# Table 1

::: {#tbl-table1 html-table-processing="true"}

```{=html}
{{< include "./Tab1.html" >}}
```

:::

## Derivation of (14)

Looking at equation (14) and the context provided, I can explain how the wage markdown is derived by combining the two markup estimates from the production approach.

### Cost minimization

From the cost minimization problem, the first-order condition for labor (equation 11) is:
$$W^l_{it} + \tilde{\lambda}_{ft} \cdot \frac{\partial W^l_{it}}{\partial L_{it}} \cdot L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \cdot \frac{P_{ft}}{\mu_{ft}}$$

Where the right side is the marginal revenue product of labor (MRPL).

Rearranging and using:
- $\beta_l$ = output elasticity of labor
- $\alpha^l_{ft} = \frac{W^l_{ft} \cdot L_{ft}}{P_{ft} \cdot Q_{ft}}$ = revenue share of labor

The markup derived from the labor first-order condition becomes:
$$\mu_{ft} = \frac{\beta_l}{\alpha^l_{ft} \cdot (1 + \tilde{\lambda}_{ft} \cdot \Psi^l)}$$

From (11)

$$W^l_{it} + \tilde{\lambda}_{ft} \frac{\partial W^l_{it}}{\partial L_{it}} L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}}$$

From the Cobb-Douglas production function: $Q_{ft} = e^{\beta_l l_{ft} + \beta_m m_{ft} + \beta_k k_{ft} + \omega_{ft}}$

Taking the derivative with respect to labor:
$$\frac{\partial Q_{ft}}{\partial L_{ft}} = \frac{\partial Q_{ft}}{\partial l_{ft}} \cdot \frac{\partial l_{ft}}{\partial L_{ft}} = \beta_l \cdot Q_{ft} \cdot \frac{1}{L_{ft}} = \beta_l \frac{Q_{ft}}{L_{ft}}$$

Substitute this into the RHS of equation (11)
$$\frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_l \frac{Q_{ft}}{L_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft} L_{ft}}$$

Multiply both sides by $L_{ft}$
$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \frac{\partial W^l_{it}}{\partial L_{it}} L_{it}^2 = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

From equation (3): $W^l_{it} = L_{it}^{\Psi^l} \eta_{it}$

Taking the derivative:
$$\frac{\partial W^l_{it}}{\partial L_{it}} = \Psi^l L_{it}^{\Psi^l - 1} \eta_{it} = \Psi^l \frac{W^l_{it}}{L_{it}}$$

Substitute this back to the above
$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \Psi^l \frac{W^l_{it}}{L_{it}} L_{it}^2 = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \Psi^l W^l_{it} L_{it} = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

$$W^l_{it} L_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l) = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

Markup
$$\mu_{ft} = \frac{\beta_l P_{ft} Q_{ft}}{W^l_{it} L_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)}$$

Define the revenue share of labor: $\alpha^l_{ft} = \frac{W^l_{ft} L_{ft}}{P_{ft} Q_{ft}}$, therefore: $W^l_{ft} L_{ft} = \alpha^l_{ft} P_{ft} Q_{ft}$

$$\mu_{ft} = \frac{\beta_l P_{ft} Q_{ft}}{\alpha^l_{ft} P_{ft} Q_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)} = \frac{\beta_l}{\alpha^l_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)}$$

This gives us the markup expression derived from the labor first-order condition, which incorporates the wage markdown term $(1 + \tilde{\lambda}_{ft} \Psi^l)$ in the denominator.

### Markup from materials (following De Loecker & Warzynski 2012)

For materials, since firms are price-takers (competitive input market), the standard markup formula applies:
$$\mu_{ft} = \frac{\beta_m}{\alpha^m_{ft}}$$

Where $\alpha^m_{ft}$ is the revenue share of materials.

From the cost minimization problem, the first-order condition for materials is:
$$W^m_{ft} = \frac{\partial Q_{ft}}{\partial M_{ft}} \frac{P_{ft}}{\mu_{ft}}$$

From the Cobb-Douglas production function: $Q_{ft} = e^{\beta_l l_{ft} + \beta_m m_{ft} + \beta_k k_{ft} + \omega_{ft}}$

Taking the derivative with respect to materials:
$$\frac{\partial Q_{ft}}{\partial M_{ft}} = \frac{\partial Q_{ft}}{\partial m_{ft}} \cdot \frac{\partial m_{ft}}{\partial M_{ft}} = \beta_m \cdot Q_{ft} \cdot \frac{1}{M_{ft}} = \beta_m \frac{Q_{ft}}{M_{ft}}$$

Substitute this into the RHS

$$W^m_{ft} = \beta_m \frac{Q_{ft}}{M_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_m \frac{P_{ft} Q_{ft}}{\mu_{ft} M_{ft}}$$

$$W^m_{ft} M_{ft} = \beta_m \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

Solve for the markup
$$\mu_{ft} = \frac{\beta_m P_{ft} Q_{ft}}{W^m_{ft} M_{ft}}$$

Define the revenue share of materials: $\alpha^m_{ft} = \frac{W^m_{ft} M_{ft}}{P_{ft} Q_{ft}}$

Therefore: $W^m_{ft} M_{ft} = \alpha^m_{ft} P_{ft} Q_{ft}$

$$\mu_{ft} = \frac{\beta_m P_{ft} Q_{ft}}{\alpha^m_{ft} P_{ft} Q_{ft}} = \frac{\beta_m}{\alpha^m_{ft}}$$

This gives us the standard De Loecker & Warzynski (2012) markup formula:
$$\mu_{ft} = \frac{\beta_m}{\alpha^m_{ft}}$$

### Equate the two markup expressions

Since both expressions equal $\mu_{ft}$:
$$\frac{\beta_l}{\alpha^l_{ft} \cdot (1 + \tilde{\lambda}_{ft} \cdot \Psi^l)} = \frac{\beta_m}{\alpha^m_{ft}}$$

Rearranging:
$$1 + \tilde{\lambda}_{ft} \cdot \Psi^l = \frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$$

From equation (12), we know that:
$$m^l_{ft} = 1 + \tilde{\lambda}_{ft} \cdot \Psi^l$$

Therefore:
$$m^l_{ft} = \frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$$

This is equation (14) in the paper.

The wage markdown $m^l_{ft}$ can be calculated in two equivalent ways:

1. **Labor supply approach**: $1 + \tilde{\lambda}_{ft} \cdot \Psi^l$ (depends on conduct parameter and labor supply elasticity)
2. **Production approach**: $\frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$ (depends only on production function parameters and cost shares)